---
title: "Exercise 09: 🔨 Text as data in R "
subtitle: "Digital disconnection on Twitter"
format: 
  html:
    toc: true
    toc-depth: 4
callout-appearance: simple
execute: 
  cache: true
  eval: false
  echo: true
  message: false
  warning: false
highlight-style: atom-one
editor_options: 
  chunk_output_type: console
---

::: {.callout-tip icon="false"}
[![Quarto Document](/img/badge-quarto_document.svg)](https://github.com/faucommsci/dbd_2023/blob/main/exercise/exercise-03.qmd) Download source file (TODO Update Link)

[![Binder RStudio](/img/badge-binder_rstudio.svg)](https://mybinder.org/v2/gh/faucommsci/dbd_binder/HEAD?urlpath=rstudio) Open this exercise in interactive and executable environment (TODO Update Binder)
:::

## Background

-   TODO Digital disconnection and why it is important
-   TODO Digital disconnection on Twitter

::: callout-note
## Todays's data basis: Twitter dataset

-   TODO Beschreibung der Erhebung:
    -   Eigene Erhebung via API
    -   Erhebungszeitraum (vor Musk Übername/Umbenennung)
    -   Search query
:::

## Packages

```{r load-packages}
# TODO Add missing packages (and description)
pacman::p_load(
  here, qs, 
  magrittr, 
  tidyverse, 
  janitor,
  easystats,
  sjmisc,
  tidytext, textdata,
  ggpubr,
  ggwordcloud)
```

## Codechunks from the session

### Import data from local

TODO Update information about central variables

| Variable            | Description                                                                    |
|:-------------------------|:---------------------------------------------|
| `movie_name`        | Name of the film                                                               |
| `release_year`      | Release year                                                                   |
| `director`          | Director of the film                                                           |
| `age_difference`    | Age difference between the characters in whole years                           |
| `couple_number`     | An identifier for the couple in case multiple couples are listed for this film |
| `actor_1_name`      | The name of the older actor in this couple                                     |
| `actor_2_name`      | The name of the younger actor in this couple                                   |
| `actor_1_birthdate` | The birthdate of the older member of the couple                                |
| `actor_2_birthdate` | The birthdate of the younger member of the couple                              |
| `actor_1_age`       | The age of the older actor when the film was released                          |
| `actor_2_age`       | The age of the younger actor when the film was released                        |

: Variablennamen und -beschreibungen

```{r import-data-raw}
# Import data from URL
tweets <- qs::qread(here("local_data/tweets-digital_detox.qs"))$raw %>% 
  janitor::clean_names()

# Check data set
tweets
```

#### Initial data check

::: callout-note
## Is the data 'technically correct'?
-   [How many records do we have? How many variables?](#check-overview)
-   [What are the variable names? Are they meaningful?](#check-overview)
-   [What type is each variable—e.g., numeric, categorical, logical?](#check-overview)
-   [How many unique values does each variable have?]
-   What value occurs most frequently, and how often does it occur?
-   Are there missing observations? If so, how frequently does this occur?

:::

##### Data overview {#check-overview}

```{r table-initial-overview}
tweets %>% glimpse()
```

##### Data processing {#check-overview-correction}

```{r recoding-data-to-correct}
tweets_correct <- tweets %>% 
  mutate(
    # reformat and create datetime variables
    across(created_at, ~ymd_hms(.)), # convert to dttm format
    year = year(created_at), 
    month = month(created_at), 
    day = day(created_at), 
    hour = hour(created_at),
    minute = minute(created_at),
    # create addtional variables
    retweet_dy = str_detect(text, "^RT"), # identify retweets
    detox_dy = str_detect(text, "#digitaldetox") 
  ) %>% 
  distinct(tweet_id, .keep_all = TRUE)
```

##### Summary statistics{#check-location-parameter}

```{r table-location-parameter}
tweets_correct %>% descr()
tweets_correct %>% skimr::skim()
```


#### Distribution over the years
<!-- TODO Add note for change of tweet character limit in November 2017 -->

```{r figure-tweets-by-year}

tweets_correct %>% 
  ggplot(aes(x = as.factor(year), fill = retweet_dy)) +
    geom_bar() +
    labs(
      x = "Year",
      y = "Number of tweets", 
      fill = "Does the tweet contain '#digitaldetox'?") +
    theme_pubr()
```

#### Top languages 
<!-- TODO Disable scientific notatino -->
```{r figure-top-languages}

tweets_correct %>% 
  group_by(lang, detox_dy) %>% 
  summarise(n = n())  %>% 
  arrange(desc(n)) %>% 
  head(n = 10) %>% 
  ggplot(aes(x = reorder(lang, n) , y = n, fill = detox_dy)) +
    geom_col() +
    labs(
      x = "Twitter language code",
      y = "Number of tweets", 
      fill = "Does the tweet contain '#digitaldetox'?") +
    coord_flip() +
    theme_pubr()
```

#### Top 'users'

```{r}
tweets_correct %>% 
  group_by(user_username, detox_dy) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n))

tweets_correct %>% 
  group_by(detox_dy) %>% 
  sjmisc::frq(
    user_username,
    min.frq = 500,
    sort.frq = "desc") 
```

### Subsample: Tweets containing #digitaldetox

```{r}
tweets_detox <- tweets_correct %>% 
  filter(
    detox_dy == TRUE, # only tweets with #digitaldetox
    retweet_dy == FALSE, # no retweets
    lang == "en" # only english tweets
    )
```

#### Summary statistics{#check-location-parameter}

```{r table-location-parameter-subsample}
tweets_detox %>% descr()
tweets_detox %>% skimr::skim()
```


## Text analysis using tidy principles
### Transform to 'tidy text'
```{r}
remove_reg <- "&amp;|&lt;|&gt;"

tidy_tweets <- tweets_detox %>% 
  mutate(text = str_remove_all(text, remove_reg)) %>% 
  tidytext::unnest_tokens("text", text) %>% 
  filter(
    !text %in% tidytext::stop_words$word)
```

### Transform to 'Summarized Text'
```{r}
summarized_tweets <- tidy_tweets %>% 
  count(text, sort = TRUE) 
```

### Visualize tokens
#### Wordcloud
```{r}
summarized_tweets %>% 
  with(wordcloud::wordcloud(text, n, max.words = 100))
```

```{r}
#| eval: false
#| echo: false

summarized_tweets %>% 
  top_n(50) %>% 
  ggplot(aes(label = text, size = n)) +
  geom_text_wordcloud() +
  scale_size_area(max_size = 30) +
  theme_minimal()
```

#### Distribution 
```{r}
summarized_tweets %>% 
  ggplot(aes(n/sum(n))) +
  geom_histogram() +
  theme_minimal()
```


## Modeling realtionships between words
### Create bigrams
```{r}
tidy_bigrams <- tweets_detox %>% 
  mutate(text = str_remove_all(text, remove_reg)) %>% 
  tidytext::unnest_tokens(
    "text", text,
    token = "ngrams", n = 2 ) %>% 
  filter(
    !text %in% tidytext::stop_words$word, 
    !is.na(text)
  )
```

### Filtering & couting
```{r}
filtered_bigrams <- tidy_bigrams %>% 
  separate(text, c("word1", "word2"), sep = " ") %>% 
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word) 
  

count_bigrams <- filtered_bigrams %>%   
  count(word1, word2, sort = TRUE)

summarized_bigrams_untied <- filtered_bigrams %>% 
  unite(bigram, word1, word2, sep = " ") %>% 
  count(bigram, sort = TRUE)
```

### Visulisation 
```{r}
library(ggraph)

bigrams_graph <- count_bigrams %>% 
  filter(n > 250) %>%
  igraph::graph_from_data_frame()

ggraph(bigrams_graph, layout = "fr") +
  geom_edge_link() +
  geom_node_point() +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1)

```

```{r}
a <- grid::arrow(type = "closed", length = unit(.15, "inches"))

ggraph(bigrams_graph, layout = "fr") +
  geom_edge_link(
    aes(edge_alpha = n), show.legend = FALSE,
    arrow = a, end_cap = circle(.07, 'inches')) +
  geom_node_point(color = "lightblue", size = 5) +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
  theme_void()
```


### Correlation

```{r}
library(widyr) 

pairs_tweets <- tidy_tweets %>% 
  pairwise_count(text, tweet_id, sort = TRUE)

word_cors <- tidy_tweets %>% 
  group_by(text) %>% 
  filter(n() >= 100) %>% 
  pairwise_cor(text, tweet_id, sort = TRUE)

word_cors %>% 
  filter(item1 %in% c("detox", "digital")) %>% 
  group_by(item1) %>% 
  slice_max(correlation, n = 5) %>% 
  ungroup() %>% 
  mutate(item2 = reorder(item2, correlation)) %>% 
  ggplot(aes(item2, correlation)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ item1, scales = "free") +
  coord_flip() +
  theme_minimal()
```



### Get sentiment 
```{r}
sentiment_tweets <- tidy_tweets %>% 
  inner_join(
     get_sentiments("bing"),
     by = c("text" = "word"),
     relationship = "many-to-many") %>% 
  count(tweet_id, sentiment) %>% 
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative) 
```

### Visualization
#### Distribution
```{r}
sentiment_tweets %>% 
  ggplot(aes(as.factor(sentiment))) +
  geom_bar() +
  theme_minimal()
```

```{r}
tweets_correct %>% 
  filter(tweet_id %in% sentiment_tweets$tweet_id) %>% 
  left_join(sentiment_tweets) %>% 
  sjmisc::rec(
    sentiment,
    rec = "-8:-1=negative; 0=neutral; 1:8=positive") %>%
  ggplot(aes(x = as.factor(year), fill = as.factor(sentiment_r))) +
    geom_bar() +
    labs(
      x = "Year",
      y = "Number of tweets", 
      fill = "Does the tweet contain '#digitaldetox'?") +
    theme_pubr()


tweets_correct %>% 
  filter(tweet_id %in% sentiment_tweets$tweet_id) %>% 
  left_join(sentiment_tweets) %>% 
  sjmisc::rec(
    sentiment,
    rec = "-8:-1=negative; 0=neutral; 1:8=positive") %>%
  ggplot(aes(x = as.factor(year), fill = as.factor(sentiment_r))) +
    geom_bar(position = "fill") +
    labs(
      x = "Year",
      y = "Number of tweets", 
      fill = "Does the tweet contain '#digitaldetox'?") +
    theme_pubr()
```

















```{r}
library(dplyr)
library(janeaustenr)
library(tidytext)

book_words <- austen_books() %>%
  unnest_tokens(word, text) %>%
  count(book, word, sort = TRUE)

total_words <- book_words %>% 
  group_by(book) %>% 
  summarize(total = sum(n))

book_words <- left_join(book_words, total_words)

book_words
#> # A tibble: 40,379 × 4
#>    book              word      n  total
#>    <fct>             <chr> <int>  <int>
#>  1 Mansfield Park    the    6206 160460
#>  2 Mansfield Park    to     5475 160460
#>  3 Mansfield Park    and    5438 160460
#>  4 Emma              to     5239 160996
#>  5 Emma              the    5201 160996
#>  6 Emma              and    4896 160996
#>  7 Mansfield Park    of     4778 160460
#>  8 Pride & Prejudice the    4331 122204
#>  9 Emma              of     4291 160996
#> 10 Pride & Prejudice to     4162 122204
#> # … with 40,369 more rows
```

```{r}
tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)

jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
```






### Filter hashtags

```{r}
tweets_detox %>% 
  mutate(hashtags = str_extract_all(text, "#\\S+")) %>% 
  unnest(hashtags) %>% 
  count(tweet_id, hashtags, sort = TRUE) 


tweets_detox %>% 
  mutate(
    user = paste0("@", user_username),
    mentions = str_extract_all(text, "@\\S+")
    ) %>% 
  unnest(mentions) %>% 
  count(mentions, sort = TRUE) 



  quanteda::fcm() %>% 
  quanteda.textplots::textplot_network()

```






### TF-IDF

```{r}
words_tweets <- tidy_tweets %>% 
  count(tweet_id, text, sort = TRUE)

words_total <- words_tweets %>% 
  group_by(tweet_id) %>% 
  summarise(total = sum(n))

tweets_words <- left_join(words_tweets, words_total)
tweets_tf_idf <- bind_tf_idf(tweets_words, text, tweet_id, n)
tweets_tf_idf %>% arrange(desc(tf_idf))
```







## 📋 Exercise: Welche Rolle spielt das Geschlecht?

::: callout-note
## Spielt das Geschlecht eine Rolle?

-   Der folgende Abschitt befasst sich nun ergänzend mit der Frage, welche Rolle das Geschlecht mit Blick auf die "Gültigkeit" der vorherigen Ergebnisse spielt
-   Dazu sind jedoch weitere Explorations- und Überarbeitungsschritte notwendig
:::

### Übeprüfung der `_gender`-Variablen

::: callout-caution
## Exercise 1

Nutzen Sie die Funktion `sjmisc::frq()` und schauen Sie sich im Datensatz `age_gaps_correct` die Variablen `actor_1_gender` und `actor_2_gender` an.
:::

```{r exercise_1}
age_gaps_correct %>% 
  frq(actor_1_gender, actor_2_gender)
```

::: callout-caution
## Exercise 2

Nutzen Sie die Funktion `sjmisc::flat_talbe()` und den Datensatz `age_gaps_correct` um eine Kreuztabelle der Variablen `actor_1_gender` und `actor_2_gender` zu erstellen.
:::

```{r exercise_2}
age_gaps_correct %>% 
  select(actor_1_gender, actor_2_gender) %>% 
  flat_table()
```

### Sind die Daten "konsistent"?

#### Überprüfung der Sortierung

```{r table-check-variable-arrangement}
age_gaps_correct %>% 
  summarise(
      p1_older = mean(actor_1_age > actor_2_age), # older person first?
      p1_male  = mean(actor_1_gender == "man"),  # male person first? 
      p_1_first_alpha = mean(actor_1_name < actor_2_name) # alphabetical order?
  )
```

### Überprüfung der Anzahl pro Paare pro Film

```{r table-check-couple-numbers}
# Create data
couples <- age_gaps_correct %>% 
  group_by(movie_name) %>% 
  summarise(n = n()) 

# Distribution
couples %>% frq(n)

# Movies with a loot of couples 
couples %>% 
  filter(n > 3) %>% 
  arrange(desc(n))
```

### Korrekturen

```{r recoding-data-to-consistent}
age_gaps_consistent <- age_gaps_correct %>% 
  # If multiple couples, assign couple number by movie
  mutate(
      couple_number = row_number(),
      .by = "movie_name"
  ) %>% 
  # Change data structure (one line per actor in a coulpe of a movie)
  pivot_longer(
    cols = starts_with(c("actor_1_", "actor_2_")),
    names_to = c(NA, NA, ".value"),
    names_sep = "_"
  ) %>% 
  # Put older actor first
  arrange(desc(age_difference), movie_name, birthdate) %>% 
    mutate(
    position = row_number(),
    .by = c("movie_name", "couple_number")
  ) %>% 
  pivot_wider(
    names_from = "position",
    names_glue = "actor_{position}_{.value}",
    values_from = c("name", "gender", "birthdate", "age")
  ) %>% 
  mutate(
    couple_structure = case_when(
      actor_1_gender == "woman" & actor_2_gender == "woman" ~ 1,
      actor_1_gender == "man" & actor_2_gender == "man" ~ 2,
      actor_1_gender != "man" ~ 3, 
      actor_1_gender == "man" ~ 4,
    ),
    older_male_hetero  = sjmisc::rec(
      couple_structure, 
      rec="3=0; 4=1; ELSE=NA", 
      to.factor = TRUE
    )
  )

```

## Die zweite Datenexploration

### Alterskombinationen im Überblick

::: callout-caution
## Exercise 3

Nutzen Sie die Funktion `sjmisc::frq()` und schauen Sie sich im Datensatz `age_gaps_consistent` die Variablen `couple_structure` und `older_male_hetero` an.
:::

```{r exercise_3}
age_gaps_consistent %>% 
  frq(couple_structure, older_male_hetero)
```

### Wie sind die Altersunterschiede unterteilt, unter Berücksichtiung des Geschlechts?

#### Graphische Überprüfung

::: callout-caution
## Exercise 4

-   Erstellen Sie, auf Basis des Datensatzes `age_gaps_consistent`, einen `ggplot`.
-   Nutzen Sie im Argument `aes()` die Variable `age_difference` als `x`-Variable und `older_male_hetero` für das Argument `fill`.
-   Nutzen Sie `geom_bar` zur Erzeugung des Plots.
-   *Optional: Verwenden Sie `theme_pubr`*
:::

```{r exercise_4}
age_gaps_consistent %>% 
  ggplot(aes(age_difference, fill = older_male_hetero)) +
  geom_bar() +
  theme_pubr()
```

```{r exercise_4_2}
age_gaps_consistent %>% 
  ggplot(aes(age_difference, fill = older_male_hetero)) +
  geom_bar() +
  labs(
    x = "Altersdifferenz (in Jahren)",
    y = 'Anzahl der "Beziehungen"'
  ) +
   scale_fill_manual(
    name = "Older partner in couple",
    values = c("0" = "#F8766D", "1" = "#00BFC4", "NA" = "grey"),
    labels = c("0" = "Woman", "1" = "Man", "NA" = "Same sex couples")
  ) +
  theme_pubr() 
```

#### Überprüfung durch Modellierung

```{r model-age-difference-release-year-gender-lm}
mdl <- lm(age_difference ~ release_year + older_male_hetero, data = age_gaps_consistent)

# Output
mdl %>% parameters::parameters()
mdl %>% performance::model_performance()
mdl %>% report::report()
```



