---
title: "Showcase 09: 🔨 Text as data in R "
subtitle: "Digital disconnection on Twitter"
format: 
  html:
    toc: true
    toc-depth: 4
callout-appearance: simple
execute: 
  cache: true
  eval: true
  echo: true
  message: false
  warning: false
highlight-style: atom-one
editor_options: 
  chunk_output_type: console
bibliography: references.bib
---

::: {.callout-tip icon="false"}
[![Quarto Document](/img/badge-quarto_document.svg)](https://github.com/faucommsci/dbd_2023/blob/main/exercise/exercise-03.qmd) Download source file (TODO Update Link)

[![Binder RStudio](/img/badge-binder_rstudio.svg)](https://mybinder.org/v2/gh/faucommsci/dbd_binder/HEAD?urlpath=rstudio) Open this exercise in interactive and executable environment (TODO Update Binder)
:::

## Background

::: {.callout-note appearance="minimal"}
## Definition

::: notes
„**Digital disconnection** is a **deliberate** (i.e., chosen by the individual) form of **non-use** of devices, platforms, features, interactions, and/or messages that occurs with higher or lower frequencies, and for shorter or longer periods of time, after the initial adoption of these technologies, and with the aim of **restoring or improving one’s perceived** overuse, social interactions, psychological **well-being**, productivity, privacy and/or perceived usefulness“. [@nassen2023]
:::
:::

-   Increasing trend towards **more conscious use of digital media (devices)**, including **(deliberate) non-use** with the aim to **restore or improve psychological well-being** (among other factors)

-   But how do "we" talk about digital detox/disconnection: 💊 drug, 👹 demon or 🍩 donut?

::: callout-important
## Todays's data basis: Twitter dataset

-   Collection of all tweets up to the beginning of 2023 that mention or discuss digital detox (and similar terms) on Twitter (not 𝕏)
-   Initial query is searching for "digital detox", "#digitaldetox", "digital_detox"
-   Access via official Academic-Twitter-API via academictwitteR [@barrie2021] at the beginning of last year
:::

## Preparation

```{r load-packages}
# TODO Add missing packages (and description)
pacman::p_load(
  here, qs, 
  magrittr, 
  tidyverse, 
  janitor,
  easystats,
  sjmisc,
  tidytext, textdata,
  ggpubr,
  ggwordcloud)
```

## Import and process the data

```{r import-data}
# Import raw data from local
tweets <- qs::qread(here("local_data/tweets-digital_detox.qs"))$raw %>% 
  janitor::clean_names()

# Initial data processing
tweets_correct <- tweets %>% 
  mutate(
    # reformat and create datetime variables
    across(created_at, ~ymd_hms(.)), # convert to dttm format
    year = year(created_at), 
    month = month(created_at), 
    day = day(created_at), 
    hour = hour(created_at),
    minute = minute(created_at),
    # create addtional variables
    retweet_dy = str_detect(text, "^RT"), # identify retweets
    detox_dy = str_detect(text, "#digitaldetox") 
  ) %>% 
  distinct(tweet_id, .keep_all = TRUE)

# Filter relevant tweets
tweets_detox <- tweets_correct %>% 
  filter(
    detox_dy == TRUE, # only tweets with #digitaldetox
    retweet_dy == FALSE, # no retweets
    lang == "en" # only english tweets
    )
```


## Chunks from the session

##### Data overview {#check-overview}

```{r table-initial-overview}
tweets %>% glimpse()
```

##### Summary statistics {#check-location-parameter}

```{r table-location-parameter}
tweets_correct %>% descr()
tweets_correct %>% skimr::skim()
```

#### Distribution over the years

<!-- TODO Add note for change of tweet character limit in November 2017 -->

```{r figure-tweets-by-year}
tweets_correct %>% 
  ggplot(aes(x = as.factor(year), fill = retweet_dy)) +
    geom_bar() +
    labs(
      x = "Year",
      y = "Number of tweets", 
      fill = "Does the tweet contain '#digitaldetox'?") +
    theme_pubr()
```

#### Top languages

<!-- TODO Disable scientific notatino -->

```{r figure-top-languages}

tweets_correct %>% 
  group_by(lang, detox_dy) %>% 
  summarise(n = n())  %>% 
  arrange(desc(n)) %>% 
  head(n = 10) %>% 
  ggplot(aes(x = reorder(lang, n) , y = n, fill = detox_dy)) +
    geom_col() +
    labs(
      x = "Twitter language code",
      y = "Number of tweets", 
      fill = "Does the tweet contain '#digitaldetox'?") +
    coord_flip() +
    theme_pubr()
```

#### Top 'users'

```{r}
tweets_correct %>% 
  group_by(user_username, detox_dy) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n))

tweets_correct %>% 
  group_by(detox_dy) %>% 
  sjmisc::frq(
    user_username,
    min.frq = 500,
    sort.frq = "desc") 
```

### Subsample: Tweets containing #digitaldetox

```{r}
tweets_detox <- tweets_correct %>% 
  filter(
    detox_dy == TRUE, # only tweets with #digitaldetox
    retweet_dy == FALSE, # no retweets
    lang == "en" # only english tweets
    )
```

#### Summary statistics

```{r table-location-parameter-subsample}
tweets_detox %>% descr()
tweets_detox %>% skimr::skim()
```

## Text analysis using tidy principles

### Transform to 'tidy text'

```{r}
remove_reg <- "&amp;|&lt;|&gt;"

tidy_tweets <- tweets_detox %>% 
  mutate(text = str_remove_all(text, remove_reg)) %>% 
  tidytext::unnest_tokens("text", text) %>% 
  filter(
    !text %in% tidytext::stop_words$word)
```

### Transform to 'Summarized Text'

```{r}
summarized_tweets <- tidy_tweets %>% 
  count(text, sort = TRUE) 
```

### Visualize tokens

#### Wordcloud

```{r}
summarized_tweets %>% 
  with(wordcloud::wordcloud(text, n, max.words = 100))
```

```{r}
#| eval: false
#| echo: false

summarized_tweets %>% 
  top_n(50) %>% 
  ggplot(aes(label = text, size = n)) +
  geom_text_wordcloud() +
  scale_size_area(max_size = 30) +
  theme_minimal()
```

#### Distribution

```{r}
summarized_tweets %>% 
  ggplot(aes(n/sum(n))) +
  geom_histogram() +
  theme_minimal()
```

## Modeling realtionships between words

### Create bigrams

```{r}
tidy_bigrams <- tweets_detox %>% 
  mutate(text = str_remove_all(text, remove_reg)) %>% 
  tidytext::unnest_tokens(
    "text", text,
    token = "ngrams", n = 2 ) %>% 
  filter(
    !text %in% tidytext::stop_words$word, 
    !is.na(text)
  )
```

### Filtering & couting

```{r}
filtered_bigrams <- tidy_bigrams %>% 
  separate(text, c("word1", "word2"), sep = " ") %>% 
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word) 
  

count_bigrams <- filtered_bigrams %>%   
  count(word1, word2, sort = TRUE)

summarized_bigrams_untied <- filtered_bigrams %>% 
  unite(bigram, word1, word2, sep = " ") %>% 
  count(bigram, sort = TRUE)
```

### Visulisation

```{r}
library(ggraph)

bigrams_graph <- count_bigrams %>% 
  filter(n > 250) %>%
  igraph::graph_from_data_frame()

ggraph(bigrams_graph, layout = "fr") +
  geom_edge_link() +
  geom_node_point() +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1)

```

```{r}
a <- grid::arrow(type = "closed", length = unit(.15, "inches"))

ggraph(bigrams_graph, layout = "fr") +
  geom_edge_link(
    aes(edge_alpha = n), show.legend = FALSE,
    arrow = a, end_cap = circle(.07, 'inches')) +
  geom_node_point(color = "lightblue", size = 5) +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
  theme_void()
```

### Correlation

```{r}
library(widyr) 

pairs_tweets <- tidy_tweets %>% 
  pairwise_count(text, tweet_id, sort = TRUE)

word_cors <- tidy_tweets %>% 
  group_by(text) %>% 
  filter(n() >= 100) %>% 
  pairwise_cor(text, tweet_id, sort = TRUE)

word_cors %>% 
  filter(item1 %in% c("detox", "digital")) %>% 
  group_by(item1) %>% 
  slice_max(correlation, n = 5) %>% 
  ungroup() %>% 
  mutate(item2 = reorder(item2, correlation)) %>% 
  ggplot(aes(item2, correlation)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ item1, scales = "free") +
  coord_flip() +
  theme_minimal()
```

### Get sentiment

```{r}
sentiment_tweets <- tidy_tweets %>% 
  inner_join(
     get_sentiments("bing"),
     by = c("text" = "word"),
     relationship = "many-to-many") %>% 
  count(tweet_id, sentiment) %>% 
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative) 
```

### Visualization

#### Distribution

```{r}
sentiment_tweets %>% 
  ggplot(aes(as.factor(sentiment))) +
  geom_bar() +
  theme_minimal()
```

```{r}
tweets_correct %>% 
  filter(tweet_id %in% sentiment_tweets$tweet_id) %>% 
  left_join(sentiment_tweets) %>% 
  sjmisc::rec(
    sentiment,
    rec = "-8:-1=negative; 0=neutral; 1:8=positive") %>%
  ggplot(aes(x = as.factor(year), fill = as.factor(sentiment_r))) +
    geom_bar() +
    labs(
      x = "Year",
      y = "Number of tweets", 
      fill = "Does the tweet contain '#digitaldetox'?") +
    theme_pubr()


tweets_correct %>% 
  filter(tweet_id %in% sentiment_tweets$tweet_id) %>% 
  left_join(sentiment_tweets) %>% 
  sjmisc::rec(
    sentiment,
    rec = "-8:-1=negative; 0=neutral; 1:8=positive") %>%
  ggplot(aes(x = as.factor(year), fill = as.factor(sentiment_r))) +
    geom_bar(position = "fill") +
    labs(
      x = "Year",
      y = "Number of tweets", 
      fill = "Does the tweet contain '#digitaldetox'?") +
    theme_pubr()
```